---
title: "The Beginnings of Tidyverse"
author: Alise Miller
format: pdf
editor: visual
---

## Task 1
### Question a reading in data 1
```{r}
library(tidyverse)

data_1 <-read_csv( "data/data.txt")
data_1
```
This has a warning, so I immediately felt like I did something wrong. Looking further into the help feature, read_csv can't be used because the data was seperated with semi-colons, maybe I could try to edit "sep = ". So this means the header is messed up along with the actual data presented. 
```{r}
data_1a <-read_csv2("data/data.txt") #makes more appropriate
data_1a
```


### Question B
```{r}
data_2 <- read_delim("data/data2.txt", 
                     col_names =TRUE, 
                     delim = "6", 
                     col_types ="fdc")

data_2

```


## Task 2
### Question a
```{r}

trailblazer <- read_csv("data/trailblazer.csv")
glimpse(trailblazer)
```

### Question b
```{r}
trailblazer_long <-
  trailblazer %>% 
   pivot_longer(cols = starts_with("game"), 
               names_to = "Location1", 
               values_to = "Points")
trailblazer_longer <- separate(trailblazer_long, 
         col= Location1, 
         into = c("Game", "Location"),
         remove = TRUE)

slice(trailblazer_longer, 1:5)

trailblazer_longer
```

### Question c 
```{r}
#trailblazer_longer |>
#  group_by(Player,Location) |>
  #summarize(mean( Points, na.rm = TRUE))
#mutate() 
#This is was what I tried to do before reading the bullet points

```

```{r}
trailblazer_wider <- trailblazer_longer |>
pivot_wider(names_from = Location, 
            values_from = Points)
trailblazer_wider

summary_of_player <- trailblazer_wider |>
  group_by(Player) |>
  summarise(mean_home = mean(Home, na.rm = TRUE),
  mean_away = mean(Away, na.rm = TRUE)
   )
summary_of_player

summary_of_player |>
  mutate(difference = mean_home - mean_away) |>
  arrange(desc(difference))

```
According to my work, the players that scored on average, more points at home than away are: Jusuf NUrkic, RObert Covington, Damian Lillard,and Cody Zeller. 

## Task 3
### Question a
```{r}
library(palmerpenguins)
penguins |>select(species, island, bill_length_mm) |>
pivot_wider(
names_from = island, values_from = bill_length_mm
)
```
The <NULL>, <dbl[52]>, and <list> mean to me that an error occured, as if the data is not formatted as my colleague intended. The <list> might indicated that the each column variables are actually lists of numbers. <dbl[52]> means entry: row Adelie and column Torgersen has 52 double or numbers with possible decimals in it. The <NULL> indicates that the entry for example entry row Chinstrap and col Torgersen is not a list or is it empty?

###Question b

```{r}
#penguins |>select(species, island, bill_length_mm) |>
  #group_by(species, island) |>
  #summarise(mean (test= bill_length_mm, na.rm = TRUE)) |>
  #pivot_wider(names_from = island, 
  #values_from = bill_length_mm)
penguins  |>
  select(island , species) |>
  count(species, island) |>
  pivot_wider(names_from = island, 
              values_from = n,
              values_fill = 0)

```

## Task 4
### Question a
```{r}
penguins |>
  select(species, bill_length_mm, island ) |>
  mutate( bill_length_mm = 
      case_when(is.na(bill_length_mm) & species == "Adelie" ~ 26, 
      is.na(bill_length_mm) & species == "Gentoo"~ 30,
      .default = bill_length_mm
  )) |>
arrange(bill_length_mm)
```


